\chapter{Decoding pedestrian and automated vehicle interactions using immersive virtual reality and interpretable deep learning}
\label{chap4}
\chaptermark{Interactions of Pedestrians and Automated Vehicles: Wait Time}
\addtocontents{toc}{\protect\setcounter{tocdepth}{1}}
\thispagestyle{empty}
\pagebreak
\section*{Preamble}
This chapter explores the features of future streets and automated vehicles' traffic parameters that may influence the  pedestrian crossing behaviour. As the market penetration rate of automated vehicles on urban streets is expected to increase in near future, they will have a significant impact on the dynamics of urban areas. Virtual reality is used to conduct hypothetical experiments on over 180 participants. Data collected is analyzed in terms of the time it takes for a pedestrian to initiate a cross in fully automated traffic, mixed traffic and human-driven traffic. Survival methods used in the previous chapter are improved with a neural network core. To enable interpretability of the results, and overcome one of the main drawbacks of machine learning methods, a game theory-based post-hoc method for model interpretability is implemented in this chapter. Suggestions on both VR data collections, and city decisions are provided based on the results of this chapter. 

\vspace{1em}
\noindent
This research article is under the third review in \textit{Transportation Research Part C: Emerging Technologies}
%\vspace{1em}

\clearpage

\section*{Abstract}
To ensure pedestrian-friendly streets in the era of automated vehicles, reassessment of current policies, practices, design, rules and regulations of urban areas is of importance. This study investigates pedestrian crossing behaviour which, as an important element of urban dynamics,
 is expected to be affected by the presence of automated vehicles. For this purpose, an interpretable machine learning framework is proposed to explore factors affecting pedestrians' wait time before crossing mid-block crosswalks in the presence of automated vehicles.
To collect rich behavioural data, we developed a dynamic and immersive virtual reality experiment, with 180 participants from a heterogeneous population in 4 different locations in the Greater Toronto Area (GTA). 
Pedestrian wait time behaviour is then analyzed using a data-driven Cox Proportional Hazards (CPH) model, in which the linear combination of the covariates is replaced by a flexible non-linear deep neural network.
The proposed model achieved a 5\% improvement in goodness of fit, but more importantly, enabled us to incorporate a richer set of covariates.
A game theoretic based interpretability method is used to understand the contribution of different covariates to the time pedestrians wait before crossing.
Results show that the presence of automated vehicles on roads, wider lane widths, high density on roads, limited sight distance, and lack of walking habits are the main contributing factors to longer wait times. Our study suggested that, to move toward pedestrian-friendly urban areas, educational programs for children, enhanced safety measures for seniors, promotion of active modes of transportation, and revised traffic rules and regulations should be considered.
\clearpage

\input{./chapter_4/section_1}

\input{./chapter_4/section_2}

\input{./chapter_4/section_3}

\input{./chapter_4/section_4}

\input{./chapter_4/section_5}

\input{./chapter_4/section_6}

\input{./chapter_4/section_7}
