\part{Dissertation Articles}
\label{part_2:dissertation_articles}
\chapter{A semi-supervised deep residual network for mode detection in Wi-Fi signals}
\label{chap2}
\chaptermark{Passive Pedestrians' Data Collection}
\addtocontents{toc}{\protect\setcounter{tocdepth}{1}}
\thispagestyle{empty}

\section*{Preamble}
This chapter introduces a deep semi-supervised residual network, trained on a passively collected Wi-Fi data, to distinguish different modes of transportation. We discuss the advantages and disadvantages of different novel data sources, and obtain a promising accuracy in detecting and monitoring different modes of transportation, including pedestrians in a congested urban environment. Counting the number of pedestrians in an area during a specific time period can be useful for tourism studies, planning purposes, forecasting required capacity, etc.

\vspace{1em}
\noindent
This research article is published in the \textit{Journal of Big Data Analytics in Transportation:}

\vspace{1em}
\noindent
Kalatian, A. and Farooq, B., 2020. A semi-supervised deep residual network for mode detection in Wi-Fi signals. Journal of Big Data Analytics in Transportation, 2(2), pp.167-180.


\vspace{1em}
\noindent

\clearpage

\section*{Abstract}
Inferring transportation mode of users in a network is of paramount importance in planning, designing, and operating intelligent transportation systems. Previous studies in the literature have mainly utilized GPS data. However, albeit the successful performances of models built upon such data, being limited to certain participants and the requirement of their involvement makes large scale implementations difficult. Due to their ubiquitous and pervasive nature, Wi-Fi networks have the potential to collect large scale, low-cost, passive and disaggregate data on multimodal transportation. In this study, by passive collection of Wi-Fi network data on a congested urban road in downtown Toronto, we attempt to tackle the aforementioned problems. We develop a semi-supervised deep residual network (ResNet) framework to utilize Wi-Fi communications obtained from smartphones. Our semi-supervised framework enables utilization of an ample amount of easily collected low-cost unlabelled data, coupled with a relatively small-sized labelled data. By incorporating a ResNet architecture as the core of the framework, we take advantage of the high-level features not considered in the traditional machine learning frameworks. The proposed framework shows a promising performance on the collected data, with a prediction precision of 81.4\% for walking, 80.5\% for biking and 84.9\% for the driving mode.
\clearpage




\input{./chapter_2/section_1}

\input{./chapter_2/section_2}

\input{./chapter_2/section_3}

\input{./chapter_2/section_4}

\input{./chapter_2/section_5}

\input{./chapter_2/section_6}