\chapter{Analysis of distracted pedestrians’ waiting time: Head-Mounted Immersive Virtual Reality application}
\label{chap3}
\chaptermark{Data Collection Using Controlled Experiments}
\addtocontents{toc}{\protect\setcounter{tocdepth}{1}}
\thispagestyle{empty}
\pagebreak
\section*{Preamble}
This chapter proposes the use of Virtual Reality (VR) as a novel source of data for assessing pedestrians' behaviours in scenarios that might be dangerous to implement in uncontrolled environments. The effect of smartphones distraction on crossing pedestrians is investigated using a survival analysis method. This study can be considered as a prelude to the studies provided in \cref{chap4} and \cref{chap6}, as the use of VR for data collection, and survival analysis for wait time are first tried and validated.

\vspace{1em} 
\noindent
This research article is published in the \textit{Journal of Collective Dynamics} as a special issue from the 2018 Conference on Pedestrian and Evacuation Dynamics: 

\vspace{1em}
\noindent
Kalatian, A., Sobhani, A. and Farooq, B., 2020. Analysis of distracted pedestrians’ waiting time: Head-Mounted Immersive Virtual Reality application. Collective Dynamics, 5, pp.46-52.
\clearpage

\section*{Abstract}
This paper analyzes the distracted pedestrians’ waiting time before crossing the road in three conditions: 1) not distracted, 2) distracted with a smartphone, and 3) distracted with a smartphone in the presence of virtual flashing LED lights on the crosswalk as a safety measure. For the means of data collection, we adapted an in-house developed virtual immersive reality environment (VIRE). A total of 42 volunteers participated in the experiment. Participants’ positions and head movements were recorded and used to calculate walking speeds, acceleration and deceleration rates, surrogate safety measures, time spent playing smartphone game, etc. After a descriptive analysis on the data, the effects of these variables on pedestrians’ waiting time are analyzed by employing a cox proportional hazards model. Several factors were identified as having impact on waiting time. The results show that an increase in initial walk speed, percentage of time the head was oriented toward smartphone during crossing, bigger minimum missed gaps and unsafe crossings resulted in shorter waiting times. On the other hand, an increase in the percentage of time the head was oriented toward smartphone during waiting time, crossing time and maze solving time, means longer waiting times for participants.
\clearpage

\input{./chapter_3/section_1}

\input{./chapter_3/section_2}

\input{./chapter_3/section_3}

\input{./chapter_3/section_4}

\input{./chapter_3/section_5}

\input{./chapter_3/section_6}

\input{./chapter_3/section_7}